Bayesian Sparsification of $\cplx$-valued networks

### 2019-11-11 paper/paper.tex
edited the cplx-vd section

% This concludes the section on the foundations of Bayesian dropout techniques.

@@

In this section we overview $\mathbb{C}$-valued networks, their motivation and peculiarities.
In our exposition we shall focus mostly on linear layers, keeping in mind that transferring 
the ideas to convolutional layers is straightforward.

The study \cite{trabelsi_deep_2017}, outlines building blocks for deep $\cplx$-valued
networks and describes suitable representation and operations including linear layers (dense
and convolutional), $\cplx$-valued activations, complex batch-normalization and suitable
weight initialization. In fact all papers touching one way or another upon the subject of
$\cplx$-valued networks, refrain each other, and the current study is no exception.

Following \cite{trabelsi_deep_2017}, we naturally identify $\cplx$ with $\real^2$ via $
  z_\mathrm{r} + j z_\mathrm{i} \mapsto (z_\mathrm{r}, z_\mathrm{i})
$, storing the real and imaginary parts in interleaved format, and emulate $\cplx$-algebra
using $\real$-valued arithmetic. In particular, convolutional and linear layers,
$
  L \colon \cplx^{\mathrm{[in]}}
    \to \cplx^{\mathrm{[out]}}
$ act upon their inputs as follows:
\begin{equation}  \label{eq:cplx-lin-op}
  \real^{2 \times \mathrm{[in]}}
    \to \real^{2 \times \mathrm{[out]}}
    \colon (x_r, x_i)
      \mapsto \Bigl(
        L_r x_r - L_i x_i,
        L_r x_i + L_i x_r
      \Bigr)
    \,,
\end{equation}
where $
  L_r, L_i
    \colon \real^{\mathrm{[in]}}
      \to \real^{\mathrm{[out]}}
$ are unique linear operators that make up $L = L_r + j L_i$. Thus, a $\cplx$-convolutional
layer with complex kernel $W$ can be implemented as two $\real$-convolutions with kernels
$\Re{W}$ and $\Im{W}$ using \eqref{eq:cplx-lin-op}. Other algebraic operations, such as
trigonometric functions, and complex division and roots in batch normalization
\cite{trabelsi_deep_2017}, are implemented through their counterparts in $\real^2$. This
makes $\cplx$-valued network into an intricately connected $\real$-valued computational
graph that respects $\cplx$-algebra.

Concerning activations, \cite{trabelsi_deep_2017} consider the following: independent
componentwise application of $\real$-valued non-linearities, $
  z = u + j v
    \mapsto \sigma(u) + j \sigma(v)
$; $\cplx$-differentiable exponential and trigonometric activations; and \textit{modRelu}
activation, which preserves the angle and applies soft-thresholding to the modulus: $
  z = r e^{j \phi}
    \mapsto e^{j \phi} (r - \theta)_+
  % z \mapsto z \bigl(
  %     1 - \tfrac\theta{\lvert z \rvert}
  %   \bigr)_+
$ with a learnable parameter $\theta\in \real$.

@@

%notation
We first clarify some notation: $M^{\top}$ is the matrix transpose, $\conj{M}$ is elementwise
complex conjugation, and $M^{\hop} = (\conj{M})^\top$ denotes Hermitian conjugate. By $e_i$ we
denote the $i$-th real unit vector of dimensionality \textit{conforming} to the matrix-vector
expression it is used in. $\vec{M}$ denotes \textit{row-major} flattening of $M$ into a vector,
i.e. in lexicographic order of its indices. $\diag{(\cdot)}$ embeds $\cplx^n$ into $n\times n$
diagonal matrices. $\otimes$ denotes the Kronecker product, for which we note the following
identities $
  \vec{P Q R} = (P \otimes R^\top) \vec{Q}
$, and $
  (P \otimes R) (C \otimes S) = P Q \otimes R S
$, \cite{petersen_matrix_2012}.
% It follows from the definition of the Kronecker product
% of $P$ and $R^\top$ and the row-major order vectorization.

@@

$\cplx$-Gaussians are closed under affine transformations:
for $A \in \cplx^{n \times m}$ and $b \in \cplx^{n}$

@@

The entropy of a $\cplx$-Gaussian distribution coincides with the entropy of
\eqref{eq:cn-paired-real-density}, and in terms of $\Gamma$ and $C$ can be expresed:

@@

In $\cplx$-variational dropout we use mean-field proper $\cplx$-Gaussian variational
approximation. Specifically, parameters $W \in \cplx^{n\times m}$ of a dense layer follow
\begin{equation}  \label{eq:c-gauss-vi}
  % (we omit the subscript $m=1$ for brevity)
  W \sim q(W)
    = \prod_{ij} q(W_{ij})
  \,,\,
    q(W_{ij})
      = \mathcal{N}^{\cplx}(W_{ij} \mid \mu_{ij}, \sigma^2_{ij}, 0)
  \,,
\end{equation}
and we use analogous independent proper $\cplx$-Gaussian approximation for convolutional
kernels. As for bias terms, we propose to use their MAP estimates, i.e. assume an independent
fully factorized Dirac approximation.

The local reparameterization trick provides both faster computations, \cite{wang_fast_2013},
and better statistical efficiency of the SGVB gradient estimator, \cite{kingma_variational_2015}.
We formulate $\cplx$ version of the local reparameterization trick using a dense linear
layer as an example, and afterwards outline the caveats for $\cplx$-convolutions.

A general mean field $\cplx$-Gaussian approximation \eqref{eq:c-gauss-vi} for $
  W \in \cplx^{n\times m}
$ is
\begin{equation}  \label{eq:c-gauss-vi-general}
  \vec{W}
    \sim \mathcal{N}^{\cplx}_{nm} \bigl(
      \vec{\mu}, \diag{\vec{\Sigma}}, \diag{\vec{C}}
    \bigr)
  \,,
\end{equation}
where mean, \textit{variance} and relation are given by matrices $
  \mu, \Sigma, C \in \cplx^{n\times m}
$ with $\Sigma_{ij} \geq 0$, and $
  \lvert C_{ij} \rvert^2 \leq \Sigma_{ij}
$. For any $x \in \cplx^m$ and $b\in \cplx^n$ $
  y = W x + b
    = (I_n \otimes x^\top) \vec{W} + b
$, whence the covariance and relation matrices of $y$ are
\begin{align}
  (I \otimes x^\top) \diag{\vec{\Sigma}} (I \otimes x^\top)^{\hop}
    % = \sum_{ij} (I \otimes x^\top) e_i \otimes e_j
    %   \Sigma_{ij} (e_i \otimes e_j)^\top (I \otimes x^\top)^{\hop}
    % = \sum_{ij} (e_i \otimes x^\top e_j)
    %   \Sigma_{ij} (e_i \otimes x^\top e_j)^{\hop}
    % = \sum_{ij} (e_i \otimes x_j) \Sigma_{ij} (e_i^\top \otimes \conj{x}_j)
    & = \sum_{i=1}^n (e_i e_i^\top) \sum_{j=1}^m \Sigma_{ij} x_j \conj{x}_j
    \,,  \label{eq:cn-gauss-lrt-cov} \\
  (I \otimes x^\top) \diag{\vec{C}} (I \otimes x^\top)^\top
    % = \sum_{ij} (e_i \otimes x_j) C_{ij} (e_i \otimes x_j)^\top
    & = \sum_{i=1}^n (e_i e_i^\top) \sum_{j=1}^m C_{ij} x_j^2
    \,.  \label{eq:cn-gauss-lrt-rel}
\end{align}
Since \eqref{eq:cn-gauss-lrt-cov} and \eqref{eq:cn-gauss-lrt-rel} are diagnal, $(y_i)_{i=1}^n$
are independent whence \eqref{eq:cn-affine} implies
\begin{equation}  \label{eq:cplx-gauss-trick}
  y_i
    \sim \mathcal{N}^{\cplx}
      \bigl(
        e_i^\top \mu x + b,
        \, \sum_{j=1}^m \Sigma_{ij} \lvert x_j \rvert^2,
        \, \sum_{j=1}^m C_{ij} x_j^2
      \bigr)
    \,.
\end{equation}
For a proper approximation \eqref{eq:c-gauss-vi} the resulting relation coefficients are zero.

\eqref{eq:cplx-gauss-trick} essentially requires three matrix-vector operations: $\mu x + b$,
$\Sigma \lvert x \rvert^2$ and $C x^2$ with the modulus and power applied elementwise. Similar
to $\real$ case, variational dropout for $\cplx$ convolutions draws independent kernel $W$
for each spatial patch in the input, since a convolution is a matrix product of $W$, unrolled
into a toeplitz matrix, and flattened intput $\vec{x}$. Weight sharing is essentially pushed
to shared variational parameters $\theta$ in the approximation $q_\theta$. The rationale
is to reduce correlation in the gradient estimator stemming from kernel weights being
shared between overlapping patches, \cite{kingma_variational_2015,citation_needed}. Thus
\eqref{eq:cplx-gauss-trick} requires computing $\cplx$ convolutions of $x$, complex square
$x^2$, and amplitude $\lvert x \rvert^2$ with the mean, relation, and variance kernels,
respectively.
%
\todo{other reasons for patch-independent weights}

@@

\begin{equation}  \label{eq:var-do-kl-div-raw}

@@

  W
    = \sum_i \lvert \theta_i + z_i \rvert^2

@@

$, i.e. non-central $\chi^2_{2m}$ with parameter $s^2$. The log-moments
of $W$ for general integer $m \geq1$ are given in \cite[p.~2466]{lapidoth_capacity_2003}.

@@

In particular, for $m=1$ and $\theta\in \cplx$ we have

@@

\begin{equation}  \label{eq:var-do-kl-div}

@@

      \tag{\ref{eq:var-do-kl-div-raw}'} \,.

@@

on \eqref{eq:var-do-kl-div}, it is possible to backpropagate through the divergence without

@@

Let's consider fully factorized proper $\cplx$-gaussian prior $
  \pi_\tau(\omega)
    \propto \mathcal{N}^{\cplx}\bigl(
      \omega \vert 0, \tau^{-1}, 0
    \bigr)
$. The per element divergence term in \eqref{eq:elbo-general-kl-div} is
\begin{equation}  \label{eq:emp-bayes-kl-div}
  KL(q \| \pi_\tau)
    = - 1 - \log{(\tau \sigma^2)}
      + \tau \bigl(
        \sigma^2 + \lvert \mu \rvert^2
      \bigr)
    \,.
\end{equation}
%
% This follows from the KL-divergence expression for two multivariate Gaussians
% $$
% % \mathbb{E}_{\omega\sim q_1(\omega)} \log \tfrac{q_1(\omega)}{q_2(\omega)}
% %   =
% KL(q_1\| q_2)
%   =
%   % - \tfrac12 \log\det{(2\pi e \Sigma_1)} + \tfrac12 \log\det{(2\pi \Sigma_2)}
%   - \tfrac12 \log \det{e I}
%   + \tfrac12 \log \frac{\det{\Sigma_2}}{\det{\Sigma_1}}
%   + \tfrac12 \tr{\bigl( \Sigma_2^{-1} \Sigma_1 \bigr)}
%   + \tfrac12 (\mu_1 - \mu_2)^\top \Sigma_2^{-1} (\mu_1 - \mu_2)
%   \,. $$
%
In empirical Bayes the prior adapts to the observed data, i.e. \eqref{eq:elbo} is optimized
w.r.t. $\tau$ of each weight's prior. Since divergence sits in \eqref{eq:elbo} with negative
sign, optimal $
  \tau^\ast = (\sigma^2 + \lvert \mu \rvert^2)^{-1}
$ is found by minimizing \eqref{eq:emp-bayes-kl-div}, which gives
\begin{equation}  \label{eq:emp-bayes-opt-kl}
  KL(q \| \pi_{\tau^\ast})
    % = - 1 - \log{({\tau^\ast} \sigma^2)}
    %   + {\tau^\ast} \sigma^2 + {\tau^\ast} \lvert \mu \rvert^2
    % = \log{((\sigma^2 + \lvert \mu \rvert^2) \tfrac1{\sigma^2})}
    = \log{\bigl(1 + \tfrac{\lvert \mu \rvert^2}{\sigma^2}\bigr)}
    % = \log{\bigl(1 + \tfrac1\alpha \bigr)}
    \,.
\end{equation}

@@

Consider the following parameterization of $W$: $
  W = \mu \odot \xi
$ where $
  \xi_{ij} \sim \mathcal{N}(1, \alpha_{ij})
$, yet $\mu \in \cplx^{n \times m}$. This case corresponds to variational inference for the
multiplicative mask $\xi$, distribution of which is approximated using independent $\real$
univariate Gaussians (ordinary dropout posteriors). Here we are actually interested in
inference, regarding the values of $\xi$ after observing the data (and fitting other model
parameters such as $\mu$). Essentially, in \eqref{eq:elbo} we have
$$
\mathbb{E}_{\xi \sim q(\xi)}
  \log p_\theta(z \vert \xi)
  - KL(q(\xi) \| \pi(\xi))
  \approx
    N \hat{\mathbb{E}}_{(z, \xi) \sim B}
      \log p_\theta(z \vert \xi)
    - KL(q(\xi) \| \pi(\xi))
  \,, $$
and by running SGVB, i.e. maximizing stochastic batched ELBO, we minimize the KL-divergence
of the variational approximation $q(\xi)$ from its true posterior $p(\xi\vert D)$. Thus we
run inference on the relevance factors of the parameters, rather than the parameters themselves.

Under this parameterization we have
$$
W_{ij} = \mu_{ij} \odot \xi_{ij}
  \sim \mathcal{N}^{\cplx} \bigl(
    \mu_{ij},
    \, \alpha_{ij} \lvert \mu_{ij} \rvert^2,
    \, \alpha_{ij} \mu_{ij}^2
  \bigr)
  \,, $$
which is a degenerate $\cplx$-Gaussian distribution. The local reparameterization trick for
this distribution implies
$$
y_i
  \sim \mathcal{N}^{\cplx}\bigl(
    e_i^\top \mu x + b,
    \, \sum_j \alpha_{ij} \lvert x_{ij} \mu_{ij}\rvert^2,
    \, \sum_j \alpha_{ij} (x_{ij} \mu_{ij})^2,
    \,
  \bigr)
  \,. $$
This means that it is a little bit more expensive to translate the local parameter noise
to the immediate activation, since we need to take into consideration the covariance between
complex output and its conjugate. Due to this, additive noise reparameterization trick,
\cite{molchanov_variational_2017}, cannot provide its variance reducing benefits, since
it is impossible to do away with dependence of the complex relation of the key parameter
$\mu_{ij}$.

The KL-divergence penalty term in this case follows form the penalty in the real variational
dropout case with the fully factorized Gaussian approximation and an improper prior $
  \pi(\xi)\propto \lvert \xi \rvert^{-\beta}
$:
$$
KL(q(\xi)\| \pi(\xi))
  % = - \tfrac12 \log 2\pi e \alpha
  %   + \tfrac\beta2 \log \alpha
  %   + \tfrac\beta2 \mathbb{E}_{z\sim \mathcal{N}(0, 1)}
  %     \log \lvert z + \tfrac1\sqrt{\alpha} \rvert^2
  \propto \tfrac{\beta-2}2 \log \alpha
    + \tfrac\beta2 \mathbb{E}_{z\sim \mathcal{N}(0, 1)}
      \log \lvert z + \tfrac1{\sqrt{\alpha}} \rvert^2
  \,. $$
Unlike the $\cplx$-valued case, even derivative of the expectation term is analytically
intractable (sec.~\ref{sub:real-chisq-grad}). Thus we use the approximation provided
by \cite{molchanov_variational_2017}.

@@

% TLDR: we use Wirtinger calculus!

>>>>

@@

In this section we briefly review $\cplx$-valued networks, their distinction from the
ordinary $\real$-valued networks.

The study \cite{trabelsi_deep_2017} outlines building blocks for deep $\cplx$-valued
networks and describes suitable representation and operations including convolutional
and dense layers, $\cplx$-valued activations, complex batch-normalization and suitable
weight initialization. Their approach makes a complex-valued network into an intricately
connected $\real$-valued computational graph that respects $\cplx$-algebra.

In particular, linear layers
$
  L \colon \cplx^{\mathrm{[in]}}
    \to \cplx^{\mathrm{[out]}}
$ act upon their inputs as follows:
\begin{equation}  \label{eq:cplx-lin-op}
  \real^{2 \times \mathrm{[in]}}
    \to \real^{2 \times \mathrm{[out]}}
    \colon (x_r, x_i)
      \mapsto \Bigl(
        L_r x_r - L_i x_i,
        L_r x_i + L_i x_r
      \Bigr)
    \,,
\end{equation}
where $
  L_r, L_i
    \colon \real^{\mathrm{[in]}}
      \to \real^{\mathrm{[out]}}
$ are unique operators that make up $L = L_r + j L_i$. A $\cplx$-convolutional layer with
kernel $W$ can be implemented as two $\real$-convolutions with kernels $\Re{W}$ and $\Im{W}$
using \eqref{eq:cplx-lin-op}. Algebraic operations, trigonometric functions, are implemented
through operations in $\real^2$. Activations, considered in \cite{trabelsi_deep_2017} include
independent componentwise application of $\real$-valued non-linearities, $
  z = u + j v
    \mapsto \sigma(u) + j \sigma(v)
$, $\cplx$-differentiable exponential and trigonometric activations, and \textit{modRelu}
activation, which applies soft-thresholding to the modulus and preserves the angle: $
  z = r e^{j \phi}
    \mapsto e^{j \phi} (r - b)_+
  % z \mapsto z \bigl(
  %     1 - \tfracb{\lvert z \rvert}
  %   \bigr)_+
$ with a learnable threshold parameter $b \in \real$.

@@

%notation
We use the following notation: $M^{\top}$ is the matrix transpose, $\conj{M}$ is elementwise
complex conjugation, and $M^{\hop} = (\conj{M})^\top$ denotes Hermitian conjugate. By $e_i$ we
denote the $i$-th real unit vector of dimensionality \textit{conforming} to the matrix-vector
expression it is used in. $\vec{M}$ denotes \textit{row-major} flattening of $M$ into a vector,
i.e. in lexicographic order of its indices. $\diag{(\cdot)}$ embeds $\cplx^n$ into $n\times n$
diagonal matrices. $\otimes$ denotes the Kronecker product, for which we note the following
identities $
  \vec{P Q R} = (P \otimes R^\top) \vec{Q}
$, and $
  (P \otimes R) (C \otimes S) = P Q \otimes R S
$, \cite{petersen_matrix_2012}.
% It follows from the definition of the Kronecker product
% of $P$ and $R^\top$ and the row-major order vectorization.

@@

$\cplx$-Gaussians are closed under affine transformations: for $A \in \cplx^{n \times m}$
and $b \in \cplx^{n}$

@@

From \eqref{eq:cn-paired-real-density} the entropy of a $\cplx$-Gaussian in terms of
$\Gamma$ and $C$ is

@@

In $\cplx$-variational dropout we use mean-field proper $\cplx$-Gaussian variational
approximation. Specifically, parameters $W \in \cplx^{n\times m}$ of a dense layer follow
\begin{equation}  \label{eq:c-gauss-vi}
  % (we omit the subscript $m=1$ for brevity)
  W \sim q(W)
    = \prod_{ij} q(W_{ij})
  \,,\,
    q(W_{ij})
      = \mathcal{N}^{\cplx}(W_{ij} \mid \mu_{ij}, \sigma^2_{ij}, 0)
  \,,
\end{equation}
and we use similar factorized approximation for convolution. We propose to MAP estimates,
of the biases i.e. assume an independent fully factorized Dirac approximation.

The local reparameterization trick provides both faster computations, \cite{wang_fast_2013},
and better statistical efficiency of the SGVB gradient estimator, \cite{kingma_variational_2015}.
A general mean field $\cplx$-Gaussian approximation \eqref{eq:c-gauss-vi} for $
  W \in \cplx^{n\times m}
$ is
\begin{equation}  \label{eq:c-gauss-vi-general}
  \vec{W}
    \sim \mathcal{N}^{\cplx}_{nm} \bigl(
      \vec{\mu}, \diag{\vec{\Sigma}}, \diag{\vec{C}}
    \bigr)
  \,,
\end{equation}
where mean, \textit{variance} and relation are given by matrices $
  \mu, \Sigma, C \in \cplx^{n\times m}
$ with $\Sigma_{ij} \geq 0$, and $
  \lvert C_{ij} \rvert^2 \leq \Sigma_{ij}
$. For any $x \in \cplx^m$ and $b\in \cplx^n$ $
  y = W x + b
    = (I_n \otimes x^\top) \vec{W} + b
$, whence the covariance and relation matrices of $y$ are
\begin{align}
  (I \otimes x^\top) \diag{\vec{\Sigma}} (I \otimes x^\top)^{\hop}
    % = \sum_{ij} (I \otimes x^\top) e_i \otimes e_j
    %   \Sigma_{ij} (e_i \otimes e_j)^\top (I \otimes x^\top)^{\hop}
    % = \sum_{ij} (e_i \otimes x^\top e_j)
    %   \Sigma_{ij} (e_i \otimes x^\top e_j)^{\hop}
    % = \sum_{ij} (e_i \otimes x_j) \Sigma_{ij} (e_i^\top \otimes \conj{x}_j)
    & = \sum_{i=1}^n (e_i e_i^\top) \sum_{j=1}^m \Sigma_{ij} x_j \conj{x}_j
    \,,  \label{eq:cn-gauss-lrt-cov} \\
  (I \otimes x^\top) \diag{\vec{C}} (I \otimes x^\top)^\top
    % = \sum_{ij} (e_i \otimes x_j) C_{ij} (e_i \otimes x_j)^\top
    & = \sum_{i=1}^n (e_i e_i^\top) \sum_{j=1}^m C_{ij} x_j^2
    \,.  \label{eq:cn-gauss-lrt-rel}
\end{align}
Since \eqref{eq:cn-gauss-lrt-cov} and \eqref{eq:cn-gauss-lrt-rel} are diagnal, $(y_i)_{i=1}^n$
are independent whence \eqref{eq:cn-affine} implies
\begin{equation}  \label{eq:cplx-gauss-trick}
  y_i
    \sim \mathcal{N}^{\cplx}
      \bigl(
        e_i^\top \mu x + b,
        \, \sum_{j=1}^m \Sigma_{ij} \lvert x_j \rvert^2,
        \, \sum_{j=1}^m C_{ij} x_j^2
      \bigr)
    \,.
\end{equation}
The trick requires three matrix operations: $\mu x + b$, $\Sigma \lvert x \rvert^2$ and
$C x^2$ with the modulus and power applied elementwise. For a proper approximation
\eqref{eq:c-gauss-vi} the relation coefficients are zero.

The trick and dropout can be applied to any layer, the output of which depends linearly
on its parameters, e.g. convolutional layers, affine and bilinear transformations ($
  y_j = x^\top W^{(j)} z + b_j
$). Similar to $\real$ case, variational dropout for $\cplx$ convolutions draws independent
kernel $W$ for each spatial patch in the input, since a convolution is a matrix product
of $W$, unrolled into a toeplitz matrix, and flattened intput $\vec{x}$. Weight sharing
is essentially pushed to shared variational parameters $\theta$ in the approximation $q_\theta$.
The rationale is to reduce correlation in the gradient estimator stemming from kernel weights
being shared between overlapping patches, \cite{kingma_variational_2015,citation_needed}.
Thus \eqref{eq:cplx-gauss-trick} requires computing $\cplx$ convolutions of $x$, complex
square $x^2$, and amplitude $\lvert x \rvert^2$ with the mean, relation, and variance kernels,
respectively.
%
\todo{other reasons for patch-independent weights}

@@

\begin{equation}  \label{eq:c-vd-kl-div-raw}

@@

  \sum_i \lvert \theta_i + z_i \rvert^2

@@

$, i.e. a non-central $\chi^2_{2m}$ with parameter $s^2$. Its log-moments for general
integer $m \geq1$ are given in \cite[p.~2466]{lapidoth_capacity_2003}. In particular,

@@

for $m=1$ and $\theta\in \cplx$ we have

@@

\begin{equation}  \label{eq:c-vd-kl-div}

@@

      \tag{\ref{eq:c-vd-kl-div-raw}'} \,.

@@

on \eqref{eq:c-vd-kl-div}, it is possible to backpropagate through the divergence without

@@

Let's consider fully factorized proper $\cplx$-gaussian prior $
  \pi_\tau(\omega)
    \propto \mathcal{N}^{\cplx}\bigl(
      \omega \vert 0, \tau^{-1}, 0
    \bigr)
$. The per element divergence term in \eqref{eq:elbo-general-kl-div} is
\begin{equation}  \label{eq:emp-bayes-kl-div}
  KL(q \| \pi_\tau)
    = - 1 - \log{(\tau \sigma^2)}
      + \tau \bigl(
        \sigma^2 + \lvert \mu \rvert^2
      \bigr)
    \,.
\end{equation}
%
% This follows from the KL-divergence expression for two multivariate Gaussians
% $$
% % \mathbb{E}_{\omega\sim q_1(\omega)} \log \tfrac{q_1(\omega)}{q_2(\omega)}
% %   =
% KL(q_1\| q_2)
%   =
%   % - \tfrac12 \log\det{(2\pi e \Sigma_1)} + \tfrac12 \log\det{(2\pi \Sigma_2)}
%   - \tfrac12 \log \det{e I}
%   + \tfrac12 \log \frac{\det{\Sigma_2}}{\det{\Sigma_1}}
%   + \tfrac12 \tr{\bigl( \Sigma_2^{-1} \Sigma_1 \bigr)}
%   + \tfrac12 (\mu_1 - \mu_2)^\top \Sigma_2^{-1} (\mu_1 - \mu_2)
%   \,. $$
%
In empirical Bayes the prior adapts to the observed data, i.e. \eqref{eq:elbo} is optimized
w.r.t. $\tau$ of each weight's prior. Since divergence sits in \eqref{eq:elbo} with negative
sign, optimal $
  \tau^\ast = (\sigma^2 + \lvert \mu \rvert^2)^{-1}
$ is found by minimizing \eqref{eq:emp-bayes-kl-div}, which gives
\begin{equation}  \label{eq:emp-bayes-opt-kl}
  KL(q \| \pi_{\tau^\ast})
    % = - 1 - \log{({\tau^\ast} \sigma^2)}
    %   + {\tau^\ast} \sigma^2 + {\tau^\ast} \lvert \mu \rvert^2
    % = \log{((\sigma^2 + \lvert \mu \rvert^2) \tfrac1{\sigma^2})}
    = \log{\bigl(1 + \tfrac{\lvert \mu \rvert^2}{\sigma^2}\bigr)}
    = \log{\bigl(1 + \tfrac1\alpha \bigr)}
    \,.
\end{equation}

@@

Consider the following parameterization of $W$: $
  W = \mu \odot \xi
$ where $
  \xi_{ij} \sim \mathcal{N}(1, \alpha_{ij})
$, yet $\mu \in \cplx^{n \times m}$. This case corresponds to inference regarding
relevance coefficients $\xi$ rather than the parameters themselves after. Under this
parameterization
\begin{equation}  \label{eq:real-scal-gauss-vi}
  W_{ij}
    \sim \mathcal{N}^{\cplx} \bigl(
      \mu_{ij},
      \, \alpha_{ij} \lvert \mu_{ij} \rvert^2,
      \, \alpha_{ij} \mu_{ij}^2
    \bigr)
  \,,
\end{equation}
which is a degenerate $\cplx$-Gaussian distribution. \eqref{eq:cplx-gauss-trick} implies
\begin{equation}  \label{eq:real-scal-gauss-trick}
  y_i
    \sim \mathcal{N}^{\cplx}\bigl(
      e_i^\top \mu x + b,
      \, \sum_j \alpha_{ij} \lvert x_{ij} \mu_{ij}\rvert^2,
      \, \sum_j \alpha_{ij} (x_{ij} \mu_{ij})^2,
      \,
    \bigr)
    \,.
\end{equation}
The KL-divergence term in \eqref{eq:elbo} coincides with \eqref{eq:improper-kl-div-real}.
We use the approximation provided by \cite{molchanov_variational_2017}. Note that additive
noise reparameterization, \cite{molchanov_variational_2017}, cannot be used in
\eqref{eq:real-scal-gauss-trick}.

@@

% TLDR: we use Wirtinger calculus!
Following \cite{trabelsi_deep_2017}, we naturally identify $\cplx$ with $\real^2$ via $
  z_\mathrm{r} + j z_\mathrm{i} \mapsto (z_\mathrm{r}, z_\mathrm{i})
$, storing the real and imaginary parts in interleaved format, and emulate $\cplx$-algebra
using $\real$-valued arithmetic.
